### 2021-09-06 tex/ms.tex
Maths

\documentclass[a4paper, 12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{dsfont}
\usepackage[left=1.5cm, right=1.5cm, bottom=2cm, top=2cm]{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage{natbib}
\newcommand{\given}{\,|\,}
\newcommand{\asin}{\sinh^{-1}}

\title{}
\author{Brendon J. Brewer}
\date{}

\begin{document}
\maketitle

%\abstract{}

% Need this after the abstract
\setlength{\parindent}{0pt}
\setlength{\parskip}{8pt}

Let the squashing function be
\begin{align}
S(x)
    &= \asin(x) \\
    &= \log\left(x + \sqrt{x^2 + 1}\right).    
\end{align}

There is a trig identity for addition:
\begin{align}
\asin u + \asin v &= \asin\left(u\sqrt{1+v^2} + v\sqrt{1+u^2}\right).
\end{align}
If we associate $u\sqrt{1+v^2}$ with the existing trending score
and $v\sqrt{1+u^2}$ with the spike mass, then the left hand side may be used
to compute the updated trending score, once we solve for $u$ and $v$.

Let $x$ be the old trending score and $m$ be the spike mass before scaling.
Then we have
\begin{align}
x &= u\sqrt{1+v^2} \\
m\exp(t/L) &= v\sqrt{1+u^2}.
\end{align}
We must now solve for $u$ and $v$ in terms of $x$ and $m$.

\end{document}

### 2021-09-09 simple_trending.py
squash/unsquash

@@ -1,37 +1,86 @@
-import bigfloat as bf
+import math
 
 # NOTE: Use deweys
 
 HALF_LIFE = 400
+LOG_TWO = math.log(2.0)
 
 def soften(lbc: float):
     assert lbc >= 0
     return lbc ** (1.0/3.0)
 
-CONTEXT = bf.Context(precision=64, emax=1000000)
-bf.setcontext(CONTEXT)
+def logsumexp(x, y):
+    top = max(x, y)
+    return top + math.log(math.exp(x - top) + math.exp(y - top))
+
+def logdiffexp(big, small):
+    assert big > small
+    return big + math.log(1.0 - math.exp(small - big))
+
+def squash(x):
+    if x < 0.0:
+        return -math.log(1.0 - x)
+    else:
+        return math.log(x + 1.0)
+
+def unsquash(x):
+    if x < 0.0:
+        return 1.0 - math.exp(-x)
+    else:
+        return math.exp(x) - 1.0
+
+
+def squashed_add(x, y):
+    """
+    squash(unsquash(x) + unsquash(y)) but avoiding overflow.
+    """
+
+    # Cases where the signs are the same
+    if x < 0.0 and y < 0.0:
+        return -logsumexp(-x, logdiffexp(-y, 0.0))
+    if x >= 0.0 and y >= 0.0:
+        return logsumexp(x, logdiffexp(y, 0.0))
+
+    # Where the signs differ
+    if x >= 0.0 and y < 0.0:
+        if abs(x) >= abs(y):
+            return logsumexp(0.0, logdiffexp(x, -y))
+        else:
+            return -logsumexp(0.0, logdiffexp(-y, x))
+    if x < 0.0 and y >= 0.0:
+        # Addition is commutative, hooray for new math
+        return squashed_add(y, x)
+
+    return None
+
+def squashed_multiply(x, y):
+    """
+    squash(unsquash(x)*unsquash(y)) but avoiding overflow.
+    """
+    sign = 1 if x*y >= 0.0 else -1
+    return sign*logsumexp(squash_to_log(abs(x)) + squash_to_log(abs(y)), 0.0)
+
+
+def log_to_squash(x):
+    return logsumexp(x, 0.0)
+
+
+def squash_to_log(x):
+    assert x > 0.0
+    return logdiffexp(x, 0.0)
+
 
 def inflate_units(height):
-    return bf.pow(bf.BigFloat(2.0), height/HALF_LIFE)
+    """
+    Log of inflated units.
+    """
[…]

### 2021-09-09 simple_trending.py
Changes to the simulation

@@ -181,18 +181,25 @@ def simulate():
     xs, ys = [], []
     xs.append(1.0)
     ys.append(claim.trending_score)
-    for height in range(2, 2000):
+    for height in range(2, 100001):
+
         if height == 20:
             claim.support_added(100.0*COIN, height)
+
+        if height == 500:
+            claim.support_abandoned(100.0*COIN, height)
+
         if height % 1000 == 0:
             claim.support_added(100.0*COIN, height)
 
         xs.append(height)
         ys.append(claim.trending_score)
+        print(height, claim.trending_score)
 
 
-    # Transform to exponential for viewing. Will blow up if simulation is too long.
+    # Transform first 10K blocks to exponential for viewing.
     xs, ys = np.array(xs), np.array(ys)
+    xs, ys = xs[0:10000], ys[0:10000]
     ys[ys >= 0.0] = np.exp(ys[ys >= 0.0]) - 1.0
     ys[ys <  0.0] = 1.0 - np.exp(-ys[ys < 0.0])
     ys /= 2.0**(xs/HALF_LIFE)

### 2021-09-09 tex/ms.tex
Maths
Rewrite squashing function

@@
 \begin{align}
-S(x)
-    &= \asin(x) \\
-    &= \log\left(x + \sqrt{x^2 + 1}\right).    
+S(x) &=
+    \left\{
+        \begin{array}{lr}
+            \log (1 + x), & \textnormal{if } x \geq 0, \\
+            -\log(1 - x), & \textnormal{otherwise}.
+        \end{array}
+    \right.
 \end{align}
-
-There is a trig identity for addition:
-\begin{align}
-\asin u + \asin v &= \asin\left(u\sqrt{1+v^2} + v\sqrt{1+u^2}\right).
-\end{align}
-If we associate $u\sqrt{1+v^2}$ with the existing trending score
-and $v\sqrt{1+u^2}$ with the spike mass, then the left hand side may be used
-to compute the updated trending score, once we solve for $u$ and $v$.
-
-Let $x$ be the old trending score and $m$ be the spike mass before scaling.
-Then we have
-\begin{align}
-x &= u\sqrt{1+v^2} \\
-m\exp(t/L) &= v\sqrt{1+u^2}.
-\end{align}
-We must now solve for $u$ and $v$ in terms of $x$ and $m$.
+Operations on the squashed grade can be defined in terms of logsumexp
+and logdiffexp, to avoid overflow.
 

### 2021-09-09 simple_trending.py
Fixed comment

@@ -1,6 +1,7 @@
 import math
 
-# NOTE: Use deweys
+# NOTE: Use deweys for amounts, except in the soften() function where the conversion
+# to LBC is assumed to have already occurred.
 
 HALF_LIFE = 400
 LOG_TWO = math.log(2.0)

### 2021-10-21 simple_trending.py
Tests for Jack

@@ -3,8 +3,8 @@ import math
 # NOTE: Use deweys for amounts, except in the soften() function where the conversion
 # to LBC is assumed to have already occurred.
 
-HALF_LIFE = 400
 LOG_TWO = math.log(2.0)
+HALF_LIFE = 399.2527760025285
 
 def soften(lbc: float):
     assert lbc >= 0
@@ -208,6 +208,15 @@ def simulate():
     plt.plot(xs, ys)
     plt.show()
 
+def test_jack():
+    claim = Claim("abc", 10*1E8, 99000)
+    print(claim.trending_score)
+    claim.claim_update(100*1E8, 99001)
+    print(claim.trending_score)
+    claim.claim_update(1000000*1E8, 99100)
+    print(claim.trending_score)
+    claim.claim_update(1*1E8, 99200)
+    print(claim.trending_score)
 
 def test_squash():
     import numpy as np

### 2021-10-21 tex/ms.tex
words

@@
 
-\title{}
+\title{The LBRY Trending Algorithm}
 \author{Brendon J. Brewer}
@@
 
-%\abstract{}
+\abstract{I describe the most recent update to the LBRY trending
+algorithm. The ideas are similar to previous implementations, but with a
+few new features.}
 
@@
 \setlength{\parskip}{8pt}
+
+\section{Introduction}
+LBRY needs a trending algorithm to rank claims by `recent popularity',
+preferably using on-chain information such as changes in support. An
+intuitive idea is to have increases and decreases in overall support cause
+a `spike' in the trending score, which then decays exponentially over time
+(assuming no further changes occur). This has formed the basis of all the
+implementations that I have written.
+
+All claims, be they streams/files, channels, reposts, or collections, have
+an associated trending score. Hubs (formerly wallet servers) compute and
+save the trending score of every claim along with the other data. Several
+views of content on LBRY Desktop and Odysee involve sorting claims by their
+trending score, in descending order.
+
+In 2021, Jack Robison rewrote a lot of the hub code to improve performance
+and to ensure hub data integrity when a chain reorganisation occurs.
+
+\section{Exponentially Decaying Spikes}
+
+\section{Nonlinearity --- the Spike Mass Function}
 

### 2021-10-21 tex/ms.tex
stuff

@@
 \usepackage{natbib}
-\newcommand{\given}{\,|\,}
-\newcommand{\asin}{\sinh^{-1}}
+\newcommand{\logsumexp}{\textnormal{logsumexp}}
 
@@
 intuitive idea is to have increases and decreases in overall support cause
-a `spike' in the trending score, which then decays exponentially over time
+a `spike' in the trending score, which then fades away over time
 (assuming no further changes occur). This has formed the basis of all the
@@
 \section{Exponentially Decaying Spikes}
+Suppose that, at time (block height) $t$, the total amount of LBC staked
+towards a claim is $\ell_t$, and its trending score is $Y_t$. At height
+$(t+1)$, if the total amount of LBC remains the same, the trending score
+is decayed towards zero:
+\begin{align}
+Y_{t+1} &= kY(t), \label{eqn:decay}
+\end{align}
+where $k$ is a decay constant which we set equal to
+$\exp(-1/576) \approx 0.9982653$, so that the e-folding timescale is 576
+blocks, approximately 24 hours.
 
-\section{Nonlinearity --- the Spike Mass Function}
+If an LBC change occurs, the update is instead
+\begin{align}
+Y_{t+1} &= kY(t) + M(\ell_t, \ell_{t+1}), \label{eqn:decay_with_spike}
+\end{align}
+where the second term is the `mass' of a `trending spike', and depends on
+the old and new LBC values. A simple idea is to set the mass of the trending
+spike equal to the change in LBC:
+\begin{align}
+M(\ell_t, \ell_{t+1}) &= \ell_{t+1} - \ell_t.
+\end{align}
+However, in practice this simple choice might lead to undesirable results.
+Many supports on the LBRY network are created by publishers themselves,
+attempting to boost their own content's visibility. If an item receives
+many organic views and a moderate number of small tips, one might expect it
+to trend, yet with the linear choice for $M$, it will have no hope of
+competing with a `whale' who can deposit 100,000 LBC in support without
+blinking an eye. In Section~\ref{sec:spike_mass}, we simulate several
+different patterns of support, and choose a function $M$ such that
+the size of trending spikes makes `minnows' able to compete with `whales',
+at least to a greater extent than they can with the linear choice.
 
-Let the squashing function be
+\section{Nonlinearity --- the Spike Mass Function}\label{sec:spike_mass}
+
+\section{Sparsity}
+For most claims, most of the time, nothing happens. However, computing the
+exponential decay using Equation~\ref{eqn:decay} would still incur a cost.
+Right now, there are about 15 million active claims on the LBRY network,
+and the intention is that there might be orders of magnitude more in the
+future. We want to avoid decaying the trending score of 15 million claims
+when nothing has happened to most of them. In the most recent implementation,
+we have found a way to avoid decaying any claims, ever --- only touched claims
+(those whose LBC values changed) ever need any computation.
+
+\section{The `Squashed' Grade}
+Suppose $x$ is a huge number, where double precision overflow might occur if
+we were to use it in a calculation (this happens at around $10^{308}$).
+If $x$ is positive, we could instead deal with logarithms. Then, to add
+two huge numbers without overflow, we could use
+\begin{align}
+\log(x + y)  &= \log(\exp(\log x) + \exp(\log y)) \\
+             &= \logsumexp(\log x, \log y),
+\end{align}
+where $\logsumexp(a, b) = $.
+
+However,
+logarithms only work for positive $x$, and our trending scores can be
+negative (and zero) as well. In principle, any monotonic function will work
+(and will allow the same total ordering of claims),
+but we need one where arithmetic is possible without overflow.
+
+Instead of $\log$, define a squashing function $S$ as:
 \begin{align}
@@
 \end{align}
-Operations on the squashed grade can be defined in terms of logsumexp
-and logdiffexp, to avoid overflow.
+The inverse, unsquashing function, is
+\begin{align}
+U(x) = S^{-1}(x) &=
+    \left\{
+        \begin{array}{lr}
+            \exp(x) - 1,  & \textnormal{if } x \geq 0, \\
+            1 - \exp(-x), & \textnormal{otherwise}.
+        \end{array}
+    \right.
+\end{align}
+
+To compute updated trending scores, we need to add values, but on the
+squashed grade. In other words, if we have squashed values $S(x)$ and
+$S(y)$, we will need the squashed sum
+\begin{align}
+S(U(x) + U(y)).
+\end{align}
+These can be worked out in terms of logsumexp and logdiffexp.
 

### 2021-11-17 tex/ms.tex
Equation
Some writing

@@
 a `spike' in the trending score, which then fades away over time
-(assuming no further changes occur). This has formed the basis of all the
+(assuming no further changes occur). This idea has formed the basis of all the
 implementations that I have written.
@@
 In 2021, Jack Robison rewrote a lot of the hub code to improve performance
-and to ensure hub data integrity when a chain reorganisation occurs.
+and to ensure hub data integrity when a chain reorganisation occurs. The
+backend now uses leveldb to store transaction and claim information, and
+Elasticsearch to store claim information in a suitable form for clients to
+query. Leveldb will be replaced with rocksdb soon.
 
@@
 \section{Nonlinearity --- the Spike Mass Function}\label{sec:spike_mass}
+As briefly discussed in the last section, a linear choice of spike mass
+function might not be ideal. When we choose this function, we are making
+a decision about how significant one amount of LBC at one time is, compared
+to a different amount of LBC at a different time. For example, perhaps
+1000 LBC right now should be considered as
+significant as 10,000 LBC 12 hours ago.
+
+Our final choice of spike mass function is complex. I will describe simpler
+versions at first, which will help in motivating the final, complicated, choice.
+\begin{align}
+M(\ell_t, \ell_{t+1}) &=
+
+\end{align}
 

### 2021-11-17 tex/ms.tex
Power

@@
 versions at first, which will help in motivating the final, complicated, choice.
+First, let's look at a choice where the LBC amounts are `softened' by raising
+them to some power less than one (here I have chosen a power of 1/3, in
+previous versions of the algorithm I used 1/4):
 \begin{align}
-M(\ell_t, \ell_{t+1}) &=
+M(\ell_t, \ell_{t+1}) &= \ell_{t+1}^{1/3} - \ell_t^{1/3}.
+\end{align}
+The size of a trending spike here would be the change in the `softened'
+LBC amount. A change from 0 to 1 LBC would give a spike mass of 1,
+as would a change from 1 to 1000 LBC, or 1000 to 1,000,000. It takes more
+LBC to move a mountain than to move a molehill.
 
+With this choice, it wouldn't matter whether a change took place in one large
+step or in smaller increments. Specifically,
+\begin{align}
+M(x, y) + M(y, z)
+    &= y^{1/3} - x^{1/3} + z^{1/3} - y^{1/3} \\
+    &= z^{1/3} - x^{1/3} \\
+    &= M(x, z).
 \end{align}
+This may not be entirely desirable. Genuinely popular claims on LBRY tend
+to receive several smaller supports, rather than one large one, and intuitively
+this ought to help. A trending algorithm should show what's trending,
+as well as what's recently been boosted.
 

### 2021-11-17 tex/ms.tex
Sparsity done

@@
 \newcommand{\logsumexp}{\textnormal{logsumexp}}
+\newcommand{\logdiffexp}{\textnormal{logdiffexp}}
 
@@
 \begin{align}
-Y_{t+1} &= kY(t), \label{eqn:decay}
+Y_{t+1} &= kY_t, \label{eqn:decay}
 \end{align}
@@
 \begin{align}
-Y_{t+1} &= kY(t) + M(\ell_t, \ell_{t+1}), \label{eqn:decay_with_spike}
+Y_{t+1} &= kY_t + M(\ell_t, \ell_{t+1}), \label{eqn:decay_with_spike}
 \end{align}
@@
 With this choice, it wouldn't matter whether a change took place in one large
-step or in smaller increments. Specifically,
+step or in smaller increments\footnote{This would occur with any
+choice of the form $M(x, y) = \Phi(y) - \Phi(x)$ for a monotonic function
+$\Phi()$.}. Specifically,
 \begin{align}
@@
 
+More to come
+
 \section{Sparsity}
@@
 when nothing has happened to most of them. In the most recent implementation,
-we have found a way to avoid decaying any claims, ever --- only touched claims
+we have devised a way to avoid decaying any claims, ever --- only touched claims
 (those whose LBC values changed) ever need any computation.
+
+The formula for updating a trending score is Equation~\ref{eqn:decay_with_spike}:
+\begin{align}
+Y_{t+1} &= kY_t + M(\ell_t, \ell_{t+1}).
+\end{align}
+However, instead of storing $Y$ for each claim, we could instead store
+a transformed version of it:
+\begin{align}
+Z_t &= Y_t k^{-t}.
+\end{align}
+The updating formula for $Y$ can then be rewritten as
+\begin{align}
+Z_{t+1} k^{t+1} &= kZ_t k^t + M(\ell_t, \ell_{t+1}).
+\end{align}
+Solving for $Z_{t+1}$ gives
+\begin{align}
+Z_{t+1} &= kZ_t k^t k^{-(t+1)} + M(\ell_t, \ell_{t+1})k^{-(t+1)} \\
+        &= Z_t + M(\ell_t, \ell_{t+1})k^{-(t+1)}.
+\end{align}
+Thus, if we maintain $Z$ instead of $Y$, 
+we replace `exponential decay' with maintenance of a constant trending
+score for all untouched claims. As time goes on, the mass of applied spikes
+merely grows exponentially (recall that $k$ is just below 1, so $k$
+raised to a negative power becomes large and positive),
+and that is how recent changes become more significant than old ones.
+
+The only difficulty now is how to add exponentially-large values to touched
+claims while avoiding overflow.
 
@@
 \end{align}
-where $\logsumexp(a, b) = $.
+where $\logsumexp$ is given by
+\begin{align}
+\logsumexp(a, b) &= \max(a, b)
+        + \log\left[\exp(a - \max(a, b)) + \exp(b - \max(a, b))\right].
+\end{align}.
+Subtracting the maximum before doing the exponentiation prevents overflow.
+Similarly, differences could be computed on the log scale. Assuming $x>y$,
+we can use the following method:
+\begin{align}
+\log(x - y)  &= \log(\exp(\log x) - \exp(\log y)) \\
+             &= \logdiffexp(\log x, \log y),
+\end{align}
+where $\logdiffexp$ is given by
+\begin{align}
+\logdiffexp(a, b) &= x
+        + \log\left[1.0 - \exp(y - x)\right].
+\end{align}.
 

### 2021-11-17 tex/ms.tex
touches

@@
 query. Leveldb will be replaced with rocksdb soon.
+As part of these changes, trending needed to be made more efficient. This
+meant we made some modifications to it, some of which are pretty neat
+in my opinion.
 
@@
 we replace `exponential decay' with maintenance of a constant trending
-score for all untouched claims. As time goes on, the mass of applied spikes
+score for all untouched claims --- the formula reduces to $Z_{t+1}=Z_t$.
+As time goes on, the mass of applied spikes
 merely grows exponentially (recall that $k$ is just below 1, so $k$
@@
 The only difficulty now is how to add exponentially-large values to touched
-claims while avoiding overflow.
+claims while avoiding overflow. This requires another nonlinear transformation
+which I informally call `squashing'.
 

### 2021-11-17 tex/ms.tex
Reference

@@
 \section{Introduction}
-LBRY needs a trending algorithm to rank claims by `recent popularity',
+LBRY \citep{li2020lbry} needs a trending algorithm to rank claims by `recent popularity',
 preferably using on-chain information such as changes in support. An
@@
 
+\bibliographystyle{plainnat}
+\bibliography{references.bib/references}
+
 \end{document}

### 2021-11-17 tex/ms.tex
Define squashed plus

@@
 \newcommand{\logdiffexp}{\textnormal{logdiffexp}}
+\newcommand{\sign}{\textnormal{sign}}
 
@@
 
-To compute updated trending scores, we need to add values, but on the
-squashed grade. In other words, if we have squashed values $S(x)$ and
-$S(y)$, we will need the squashed sum
+To compute updated trending scores, we need to add and multiply
+values, but on the
+squashed grade. In other words, if we have squashed values $x$ and
+$y$, we will need the squashed sum
 \begin{align}
-S(U(x) + U(y)).
+x \oplus y &= S(U(x) + U(y))
+\end{align}
+and the squashed product
+\begin{align}
+x \otimes y &= S(U(x)U(y)).
 \end{align}
 These can be worked out in terms of logsumexp and logdiffexp.
+The results are fairly complicated and are given in the Appendix.
 
@@
 
+\appendix
+\section{Squashed Arithmetic}
+The squashed addition operator is broken down into four cases, depending
+on the sign of the arguments:
+\begin{align}
+x \oplus y &= \left\{
+    \begin{array}{lr}
+   -\logsumexp(-x, \logdiffexp(-y, 0)),& x < 0~\textnormal{and}~y < 0, \\
+    \logsumexp(x, \logdiffexp(y, 0)),& x \geq 0~\textnormal{and}~y \geq 0 \\
+    \logsumexp(0, \logdiffexp(x, -y)), & \sign(x) \neq \sign(y)~\textnormal{and}~|x| \geq |y|\\
+    -\logsumexp(0, \logdiffexp(-y, x)), & \sign(x) \neq \sign(y)~\textnormal{and}~|x| < |y|
+    \end{array}
+\right. .
+\end{align}
+
 \end{document}

### 2021-11-18 tex/ms.tex
Spike power

@@
 
+\subsection{Change in Softened LBC Amount}
 Our final choice of spike mass function is complex. I will describe simpler
@@
 \begin{align}
-M(\ell_t, \ell_{t+1}) &= \ell_{t+1}^{1/3} - \ell_t^{1/3}.
+M_1(\ell_t, \ell_{t+1}) &= \ell_{t+1}^{1/3} - \ell_t^{1/3}.
 \end{align}
@@
 \begin{align}
-M(x, y) + M(y, z)
+M_1(x, y) + M_1(y, z)
     &= y^{1/3} - x^{1/3} + z^{1/3} - y^{1/3} \\
     &= z^{1/3} - x^{1/3} \\
-    &= M(x, z).
+    &= M_1(x, z).
 \end{align}
@@
 
-More to come
+\subsection{Softened Change in LBC Amount}
+Instead of applying the power before the subtraction, we could apply the
+subtraction first --- i.e., we could use
+\begin{align}
+M_2(\ell_t, \ell_{t+1}) &= \left(\ell_{t+1} - \ell_t\right)^{1/3}.
+\end{align}
+The subscript 2 distinguishes this choice from $M_1()$ in the previous
+subsection.
+This would make several small supports more effective than one large support,
+in terms of the trending score, a potentially desirable feature:
+\begin{align}
+M_2(x, y) + M_2(y, z)
+    &= (y - x)^{1/3} + (z - y)^{1/3} \\
+    &> (z - x)^{1/3}.
+\end{align}
+The inequality follows from (I don't know --- Jensen's inequality perhaps).
+However, in numerical experiments, we found that this effect was too strong,
+and whales could attain huge heights in trending score by splitting up, say,
+100,000 LBC, into several supports.
+
+\subsection{Interpolation}
+The two ideas $M_1()$ and $M_2()$ can be merged into a compromise third option,
+as follows:
+\begin{align}
+M_3(\ell_t, \ell_{t+1}) &= M_1(\ell_t, \ell_{t+1})^\alpha
+                           M_2(\ell_t, \ell_{t+1})^{1 - \alpha}.
+\end{align}
+This is linear interpolation of $\log M$ values, or alternatively, it is
+`geometric interpolation'. If $\alpha$ is close to 1 then the rule is more
+like $M_1$, if it is close to zero it is more like $M_2$.
+
+What we actually settled on was to have $\alpha \approx 0.5$ when we're dealing
+with low LBC values, and $\alpha = 0.85$ when dealing with higher LBC values.
+There is a narrow range over which $\alpha$ is between these two values:
+\begin{align}
+\alpha &= \left\{
+        \begin{array}{lr}
+            0.5, & \ell_2 \leq 50,\\
+            0.01\ell_2, & 50 < \ell_2 \leq 85,\\
+            0.85, & \textnormal{otherwise}.
+        \end{array}
+        \right.
+\end{align}
+This completes the specification of the spike mass function.
 

### 2021-11-18 tex/ms.tex
More sections written

@@
 M_3(\ell_t, \ell_{t+1}) &= M_1(\ell_t, \ell_{t+1})^\alpha
-                           M_2(\ell_t, \ell_{t+1})^{1 - \alpha}.
+                           M_2(\ell_t, \ell_{t+1})^{1 - \alpha}. \label{eqn:spike_mass}
 \end{align}
@@
 This completes the specification of the spike mass function.
+
+% TODO: Consider idea of symmetrising by using both old and new LBC values,
+% as this may make reorg safety occur naturally without any extra effort
 
@@
 Z_{t+1} &= kZ_t k^t k^{-(t+1)} + M(\ell_t, \ell_{t+1})k^{-(t+1)} \\
-        &= Z_t + M(\ell_t, \ell_{t+1})k^{-(t+1)}.
+        &= Z_t + M(\ell_t, \ell_{t+1})k^{-(t+1)}. \label{eqn:sparse_update}
 \end{align}
@@
 \end{align}
-where $\logsumexp$ is given by
+where $\logsumexp()$ is given by
 \begin{align}
@@
         + \log\left[\exp(a - \max(a, b)) + \exp(b - \max(a, b))\right].
-\end{align}.
+\end{align}
 Subtracting the maximum before doing the exponentiation prevents overflow.
@@
 \begin{align}
-\logdiffexp(a, b) &= x
-        + \log\left[1.0 - \exp(y - x)\right].
-\end{align}.
-
+\logdiffexp(a, b) &= a
+        + \log\left[a - \exp(b-a)\right].
+\end{align}
 However,
@@
 \end{align}
+For large positive values, these are approximately equal to log and exp
+respectively, but for smaller values (and negative values), it remains
+continuous.
 
 To compute updated trending scores, we need to add and multiply
-values, but on the
+values using Equations~\ref{eqn:sparse_update}~and~\ref{eqn:spike_mass}, but on the
 squashed grade. In other words, if we have squashed values $x$ and
@@
 
+Squashed multiplication is simpler:
+\begin{align}
+x \otimes y &= \sign(xy)
+                \logsumexp\left(
+                    \logdiffexp(|x|, 0) + \logdiffexp(|y|, 0),
+                0\right).
+\end{align}
+
 \end{document}

### 2021-11-18 tex/ms.tex
Fixed logdiffexp
Kevin

@@
 \logdiffexp(a, b) &= a
-        + \log\left[a - \exp(b-a)\right].
+        + \log\left[1 - \exp(b-a)\right]
 \end{align}
+for $a > b$.
+
 However,
@@
 
+\section{Acknowledgements}
+BJB would like to thank Kevin Knuth (SUNY Albany) for inspiration.
+
 \bibliographystyle{plainnat}

### 2021-11-18 tex/ms.tex
Big numbers

@@
 
+In the actual code, we store squashed trending values, i.e., $S(Z)$ in the
+notation of this paper, for every claim. When claims are touched, the
+squashed trending values are updated. At block height 1,000,000, which occurred
+in 2021, unsquashed trending values would have been of order
+\begin{align}
+\exp\left(\frac{10^{6}}{576}\right) &\approx 10^{753}.
+\end{align}
+However, the squashed values are only of order $10^6 / 576 \approx 1736$.
+Differences in squashed trending values, which determine the ordering of claims
+in applications, occur in the third, fourth, and subsequent digits.
+With double precision storage, the situation will remain acceptable for
+centuries. For example, at block height 100 million, expected around 2496 AD,
+squashed trending scores will be around 150,000, with plenty of remaining
+significant digits to allow for sorting.
+
 \section{Acknowledgements}

### 2021-11-18 tex/ms.tex
Sentence fixup

@@
 \end{align}
-These can be worked out in terms of logsumexp and logdiffexp.
-The results are fairly complicated and are given in the Appendix.
+These can be worked out in terms of logsumexp and logdiffexp, and thus can
+be computed relatively easily without overflow.
+The results are given in the Appendix.
 

### 2021-11-18 tex/ms.tex
Figure

@@
 \end{align}
+For a claim created at height $t=50$ with 1 LBC staked towards it, and then
+later supported with 3 more LBC at height $t=550$, the trending score
+would respond as in Figure~\ref{fig:simple}.
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=0.7\textwidth]{figures/simple.pdf}
+\caption{An example of how the trending score might vary over time
+for a single claim.\label{fig:simple}}
+\end{figure}
+
 However, in practice this simple choice might lead to undesirable results.

### 2021-11-18 tex/simple.py

@@ -1,0 +1,20 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+# Fonts for plot
+plt.rc("font", size=14, family="serif", serif="Computer Sans")
+plt.rc("text", usetex=True)
+
+ts = np.arange(1, 1001)
+kernel1 = np.zeros(len(ts))
+kernel1[ts >= 50] += np.exp(-ts[ts >= 50]/576)
+kernel1 = kernel1/np.max(kernel1)
+kernel2 = np.zeros(len(ts))
+kernel2[ts >= 550] += np.exp(-ts[ts >= 550]/576)
+kernel2 = 3*kernel2/np.max(kernel2)
+ys = kernel1 + kernel2
+
+plt.plot(ts, ys, "-")
+plt.xlabel("Block Height $t$")
+plt.ylabel("Trending Score $Y$")
+plt.savefig("figures/simple.pdf")

### 2021-11-18 tex/ms.tex
New figure

@@
 respectively, but for smaller values (and negative values), it remains
-continuous.
+continuous --- see Figure~{fig:squash}.
+
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=0.7\textwidth]{figures/squash.pdf}
+\caption{The squashing function $S(x)$ is the solid black curve.
+The dashed blue and pink curves are $\log(x)$ and $-\log(-x)$ respectively.
+\label{fig:squash}}
+\end{figure}
 

### 2021-11-18 tex/squash.py

@@ -1,0 +1,20 @@
+import matplotlib.pyplot as plt
+import numpy as np
+
+# Fonts for plot
+plt.rc("font", size=14, family="serif", serif="Computer Sans")
+plt.rc("text", usetex=True)
+
+xs = np.linspace(-100.0, 100.0, 10001)
+ys = np.empty(len(xs))
+rhs = xs >= 0.0
+ys[rhs] = np.log(xs[rhs] + 1.0)
+ys[~rhs] = -np.log(1.0 - xs[~rhs])
+plt.plot(xs, ys, "k-", label="Squashed $S(x)$")
+plt.plot(xs[xs > 0.0], np.log(xs[xs > 0.0]), "b--", alpha=0.5, label="$\log(x)$")
+plt.plot(xs[xs < 0.0], -np.log(-xs[xs < 0.0]), "r--", alpha=0.5, label="-$\log(-x)$")
+plt.xlabel("Value $x$")
+plt.ylabel("Transformed Value")
+plt.legend()
+plt.savefig("figures/squash.pdf")
+

### 2021-11-28 simple_trending.py
Trending

@@ -167,6 +167,7 @@ class Claim:
 
     def __str__(self):
         return str(dict(claim_hash=self.claim_hash,
+                        bid=self.bid,
                         support_amount=self.support_amount,
                         trending_score=self.trending_score))
 

### 2021-11-28 tex/response.py

@@ -1,0 +1,48 @@
+import matplotlib.pyplot as plt
+import numpy as np
+import sys
+sys.path.append("../")
+from simple_trending import *
+
+# Fonts for plot
+plt.rc("font", size=14, family="serif", serif="Computer Sans")
+plt.rc("text", usetex=True)
+
+COIN = 1E8
+
+claim = Claim("abc", 1.0*COIN, 1)
+xs, ys = [], []
+xs.append(1.0)
+ys.append(claim.trending_score)
+for height in range(2, 1000):
+
+    if height == 20:
+        claim.support_added(100.0*COIN, height)
+
+    if height == 500:
+        claim.support_added(1000.0*COIN, height)
+
+    if height == 700:
+        claim.support_abandoned(1000.0*COIN, height)
+
+    xs.append(height)
+    ys.append(claim.trending_score)
+
+
+plt.plot(xs, ys, label="Squashed $S(Z)$")
+
+# Transform first 10K blocks to exponential for viewing.
+xs, ys = np.array(xs), np.array(ys)
+ys[ys >= 0.0] = np.exp(ys[ys >= 0.0]) - 1.0
+ys[ys <  0.0] = 1.0 - np.exp(-ys[ys < 0.0])
+
+plt.plot(xs, ys, label="Sparse $Z$")
+ys /= 2.0**(xs/HALF_LIFE)
+
+plt.plot(xs, ys, label="Original $Y$")
+
+plt.xlabel("Time $t$ (blocks)")
+plt.ylabel("Trending Score")
+plt.legend()
+plt.show()
+

### 2021-11-28 tex/response.py
Graph tweaks

@@ -14,15 +14,15 @@ claim = Claim("abc", 1.0*COIN, 1)
 xs, ys = [], []
 xs.append(1.0)
 ys.append(claim.trending_score)
-for height in range(2, 1000):
+for height in range(2, 5000):
 
-    if height == 20:
+    if height == 200:
         claim.support_added(100.0*COIN, height)
 
-    if height == 500:
+    if height == 1200:
         claim.support_added(1000.0*COIN, height)
 
-    if height == 700:
+    if height == 1500:
         claim.support_abandoned(1000.0*COIN, height)
 
     xs.append(height)

### 2021-11-28 tex/response.py
Figure stuff

@@ -44,5 +44,4 @@ plt.plot(xs, ys, label="Original $Y$")
 plt.xlabel("Time $t$ (blocks)")
 plt.ylabel("Trending Score")
 plt.legend()
-plt.show()
-
+plt.savefig("figures/response.pdf")

### 2021-11-28 tex/squash.py

@@ -12,7 +12,7 @@ ys[rhs] = np.log(xs[rhs] + 1.0)
 ys[~rhs] = -np.log(1.0 - xs[~rhs])
 plt.plot(xs, ys, "k-", label="Squashed $S(x)$")
 plt.plot(xs[xs > 0.0], np.log(xs[xs > 0.0]), "b--", alpha=0.5, label="$\log(x)$")
-plt.plot(xs[xs < 0.0], -np.log(-xs[xs < 0.0]), "r--", alpha=0.5, label="-$\log(-x)$")
+plt.plot(xs[xs < 0.0], -np.log(-xs[xs < 0.0]), "r--", alpha=0.5, label="$-\log(-x)$")
 plt.xlabel("Value $x$")
 plt.ylabel("Transformed Value")
 plt.legend()

### 2021-12-15 tex/ms.tex
Figure stuff
More writing

@@
 \title{The LBRY Trending Algorithm}
-\author{Brendon J. Brewer}
+\author{Brendon J. Brewer and Jack Robison}
 \date{}
@@
 a `spike' in the trending score, which then fades away over time
-(assuming no further changes occur). This idea has formed the basis of all the
-implementations that I have written.
+(assuming no further changes occur). This idea, which corresponds to
+AR(1) time series in statistics \citep{brockwell2009time}
+has formed the basis of all the implementations I have written --- though
+other implementations have also existed in the past.
 
@@
 In 2021, Jack Robison rewrote a lot of the hub code to improve performance
-and to ensure hub data integrity when a chain reorganisation occurs. The
+and to ensure hub data integrity when a blockchain reorganisation occurs. The
 backend now uses leveldb to store transaction and claim information, and
@@
 
-% TODO: Consider idea of symmetrising by using both old and new LBC values,
-% as this may make reorg safety occur naturally without any extra effort
+{\bf TODO}: Consider idea of symmetrising by using both old and new LBC values,
+as this may make reorg safety occur naturally without any extra effort
+
+{\bf TODO}: Show the response of the trending score to different patterns
+of support.
+
+{\bf TODO}: Bring time delays back in, and write about why.
 
@@
 for $a > b$.
+If we were to use logarithms, all orderings would remain the same, since
+log is a monotonic function \citep{knuth2019lattices}.
 
@@
 
+\begin{figure}[!ht]
+\centering
+\includegraphics[width=0.7\textwidth]{figures/response.pdf}
+\caption{An example of how the trending score might vary over time
+for a single claim. The green curve shows $Y(t)$ which illustrates
+the `exponentially decaying spikes' concept. The orange curve
+shows $Z(t)$ which uses the sparse representation, so the trending
+score remains constant whenever nothing happens to the claim.
+Finally, the blue curve shows $S(Z(t))$, after the squashing
+function is applied to prevent overflow. This is what the hub server
+actually computes.\label{fig:response}}
+\end{figure}
+
 In the actual code, we store squashed trending values, i.e., $S(Z)$ in the
@@
 \section{Acknowledgements}
-BJB would like to thank Kevin Knuth (SUNY Albany) for inspiration.
+BJB would like to thank Kevin Knuth (SUNY Albany) for inspiration, and the
+LBRY Inc staff for building such a cool thing.
 
